---
title: 'BIOL-812 #1st Assignment'
author: "Andrew Le"
date: '2022-03-26'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install Packages
```{r}
install.packages("TPD")
install.packages("ks")
install.packages("dplyr")
install.packages("ggfortify")
install.packages("ggplot2")
install.packages("cowplot")
install.packages("missMDA")
install.packages("FactoMineR")
```
Load Library
```{r}
library(TPD)
library(ks)
library(dplyr)
library(ggfortify)
library(ggplot2)
library(cowplot)
library(missMDA)
library(FactoMineR)
```

Import Data
```{r}
setwd("/Users/andrewle//Desktop/BIOL812") 
traitdryaddata<- read.csv("traits.ind 1.csv")
matrixfreqdryaddata<-read.csv("matrix.freq 1.csv")
```

View Data
```{r}
View(traitdryaddata)
View(matrixfreqdryaddata)
```

Check for NA and amount of NA in the Trait dataframe
```{r}
is.na(traitdryaddata)
sum(is.na(traitdryaddata))
```

Check for NA and amount of NA in the Frequency dataframe
```{r}
is.na(matrixfreqdryaddata)
sum(is.na(matrixfreqdryaddata))
```

Size-Correction for Trait Data with respect to body size
```{r}
View(traitdryaddata)
traitdryaddata1<-traitdryaddata
traitdryaddata1[,3:8]<-traitdryaddata[,3:8]/traitdryaddata[,2]
traitdryaddata2<-traitdryaddata1 %>% select(sp,head.width:leg.length)
View(traitdryaddata2)
```

PCA:log transformation
https://www.geeksforgeeks.org/how-to-make-pca-plot-with-r/
```{r}
log.trait<-log(traitdryaddata2[,2:7])
trait.species<-traitdryaddata2[,1]
```

PCA: Results (Omitting Row(s) with N/A)
```{r}
trait.pca<-prcomp(na.omit(log.trait),center=TRUE,scale. = TRUE)
print(trait.pca)
summary(trait.pca)
```
PCA: Plot for Variation, to see which PC accounts for the most variation
```{r}
plot(trait.pca,type = "l")
```
PCA for Component 1 and 2
```{r}
autoplot(trait.pca, data=na.omit(traitdryaddata2), colour="sp")
```
Biplot to see which trait accounted for what variation
```{r}
biplot.trait.pca<-biplot(trait.pca,xlim=c(-0.20,0.20) ,cex=.5);biplot.trait.pca
```
Most variation found in PC1, but doesn't agree with paper
Paper: 
PC1 (55.1%):Negatively head and pronotum width | Positively Leg Length
PC2 (21.6%):Negatively eye width | Positively Body size

Our PCA:
PC1 (52.84%): Negatively leg length and scape length | Positively Mandible Length and Pronotum and Head Width
PC2 (23.34%): Negatively eye width


PCA Summary and Print (removing certain values)
```{r}
nb<-estim_ncpPCA(log.trait,scale=TRUE)
missMDAPCA<- imputePCA(log.trait, ncp=2, scale = TRUE)
misstraitPCA<-PCA(missMDAPCA$completeObs)
print(missMDAPCA)
summary(missMDAPCA)
```
Still difference in Variation from Paper
PC1 = 51.24
PC2 = 24.78


Trait Probability Density (TPD)
```{r}
naomittraitdata<-na.omit(traitdryaddata2)

par(mfrow=(c(2,3)))
#Head Width
trait_headwidth<-density(naomittraitdata$head.width)
THW<-plot(trait_headwidth, main = "TPD Head Width")

#Pronotum Width
trait_pronotumwidth<-density(naomittraitdata$pronotum.width)
TPW<-plot(trait_pronotumwidth, main = "TPD Pronotum Width")

#Mandible Length
trait_mandiblelength<-density(naomittraitdata$mandible.length)
TML<-plot(trait_mandiblelength, main = "TPD Mandible Length")

#Scape Length
trait_scapelength<-density(naomittraitdata$scape.length)
TSL<-plot(trait_scapelength, main = "TPD Scape Length")

#Eye Width
trait_eyewidth<-density(naomittraitdata$eye.width)
TEW<-plot(trait_eyewidth, main = "TPD Eye Width")

#Leg Length
trait_leglength<-density(naomittraitdata$leg.length)
TLL<-plot(trait_leglength, main = "TPD Leg Length")
```

TPD for two Traits (Head Width and Mandible Length)
```{r}
traits_matrix<-naomittraitdata[,c("head.width","mandible.length")]
sp_matrix<-naomittraitdata$sp
TPDs_result<-TPDs(species = sp_matrix,traits_matrix)
```

Plot TPD 
```{r}
plotTPD(TPD=TPDs_result)
```
X-axis is head width
Y-axis is mandible length
Lighter tones of blue shows higher probability of trait combination
Darker tones of blue shows lower probaility of trait combination


Alpha determines the proportion of the probability density function of each species or population that will be included in the resulting TPD
```{r}
traits_matrix_d1<-traits_matrix[,c("head.width")]
TPDs_result_d1_a1<-TPDs(species = sp_matrix, traits=traits_matrix_d1,alpha=0.95)
plotTPD(TPD = TPDs_result_d1_a1)
```

Estimating Functional Similarities and Dissimilarity (first six species)
```{r}
natrait<-naomittraitdata
natrait<-as.data.frame(unclass(natrait),stringsAsFactors = TRUE)
firstSix<-droplevels(unique(natrait$sp)[1:6])
traitsSix<-subset(natrait,natrait$sp %in% firstSix)
TPDsSix<-TPDs(species=droplevels(traitsSix$sp), traits = traitsSix$head.width, alpha =0.95)
```
Plotting TPD of Height
```{r}
par(mfrow=c(1,1))
plot(TPDsSix$data$evaluation_grid,TPDsSix$TPDs[[1]], type = "n", ylim=c(0,0.02))
for(i in 1:length(TPDsSix$TPDs)){
  lines(TPDsSix$data$evaluation_grid, TPDsSix$TPDs[[i]], lwd=2, col=i)
}
legend("topright", bty = "n", lwd = 2, col = 1:length(TPDsSix$TPDs),legend=names(TPDsSix$TPDs[1:6]))
```
No Species overlap


Dissimilarities 
```{r}
dissimSix<- dissim(TPDsSix)
dissimSix$populations$dissimilarity
```






